\section{Introduction} \label{sect:intro}

\subsection{Background and Motivation} \label{subsect:intro:background}

% What is VASCO?
% Why is such problem important to study?
The ``Vanishing and Appearing Sources during a Century of Observations'' (\vasco) initiative aims at finding inexplicable effects among all-sky surveys. The \vasco project is a collaboration between astronomers and information technology researchers, and incorporates explicitly a component of citizen science. The study of differences among all-sky surveys could lead to interesting scientific findings, like new astrophysical phenomena or interesting targets for follow-up by the Search for Extraterrestrial Intelligence (\seti) observations.

% What is the problem?
The enormous scale of existing sky surveys motivates the development of new and efficient computational tools, with an exciting role given to machine learning (\ml) due to its capacity to deal with data-intensive processes.

% What is the goal?
The precise objective of this project is to implement and test an \ml algorithm which uses a data-driven approach to attempt to learn what features describe interesting candidates from the historical sky survey observations. The \ml component is described as \mlblink and it is based on methods of active and online semi-supervised learning. \mlblink is named after the blink comparator; a 19th century viewing device invented by physicist Carl Pulfrich used by astronomers to discover differences between two images of the night sky.

\subsection{Recommender Systems} \label{subsect:intro:recommender-systems}

% What is a recommender system?
A recommender system is a computer software which allows to provide product suggestions that serve a certain purpose to an entity. The entity to which such recommendation is provided is usually referred to as the user, while the product being recommended is commonly referred to as an item \cite{book:rs}.

% Why recommender systems?
% What is the goal of a recommender system?
The usage of a recommender system is typically motivated by the existence of a set of predefined objectives to optimize and a possibly overwhelming number of items to choose from. A recommender system's goal is to maximize the established set of objectives; a goal which can be accomplished by the use of a data--driven approach which attempts to learn existing dependencies among users and items.

% How does a simple recommender system work?
As an example, consider an online bookstore that uses a recommender system to suggest books to its users. Such system might utilize explicit feedback such as a star rating system (e.g., 0--5), or implicit feedback like browsing for a title or buying a book to infer its users interests. The recommender system prediction based off the data aforementioned can then be used to increase profit and user engagement in the platform.

\subsection{Outline of Thesis} \label{subsect:intro:outline}
